A Survey of Scientific Workflow Management Systems

### 2014-04-06 paper.tex
drop GWES

%Workflows considered:
% MOTEUR2
% Taverna
% Swift
% Pegasus
% Triana
% Galaxy
% Makeflow
% Dryad
% Science pipes
% GWES
% Spark
% Linda
% DaGue

@@

\section{Objectives and Organization}
The present work analyzes contemporary SWEs used in the context of e-science to
find out the capabilities and key features in terms of the characteristics
described in the introduction part of this paper. The rest of this paper is
organized as follows: Section \ref{sec:overview} presents an overview of the
key SWEs studied in this work. Section \ref{sec:scal} discusses the
optimization features and resulting scalable performance as observed in the
SWEs studied. Section \ref{sec:data} studies the data management and
description aspects of the SWEs. Section \ref{sec:inter} presents a study of
SWEs approaches to interface the large-scale computing infrastructures. Section
\ref{sec:lang} details the language features and expressiveness of the SWEs
studied. Section \ref{sec:usable} discusses the usability aspects of the SWEs.
Section \ref{sec:sum} presents a summary and a classification of all the
aspects of SWEs studied in this work. Throughout the paper, we also present
tables summarizing the aspects of SWEs considered in respective sections. It
was not possible to obtain an exhaustive information about all the SWEs
considered in this work since we rely on the peer reviewed publications for
information on a given SWE. Some features, such as extendability of SWE
strongly depend on the extension points designed in the code; consequently, we
consider it present in only those SWEs that have a documented description or
demonstration of the given feature.

@@

% Put this somewhere else, either in the analysis or may be in the conclusions and summary.
%The \textit{gscript} \cite{maheshwari-montagnat:2010} workflow language developed as a part of the current work is similar in spirit to the SwiftScript. Similar to SwiftScript, gscript adopts futures to express asynchronous execution in a script-based language. A SwiftScript workflow contains statically typed datasets that can be specified using ``annotations'' in the script. Unlike SwiftScript, gscript does provide user-defined strategies in order to effect data combinations in the workflow. SwiftScript provides for fault-tolerance and exception handling at run-time which is lacking currently in gscript. The underlying datasets organization in swift is represented using an XDTM XML-representation. This means that effectively, a workflow is SwiftScript plus the XDTM data representation. A workflow is simpler and compact in gscript as its data description is embedded within a workflow representation. SwiftScript supports creation of nested-workflows in the form of compound procedures. Conditional and Loop execution is supported using the if and for/while statements in both the languages.

@@

%
\paragraph{GWES/GWorkflowDL} The Grid Workflow Execution Service
(GWES)~\cite{Hoheisel2007} is a SWE developed by the Fraunhofer-First
institute. It implements a dynamic workflow composition and enactment model
based on the Grid Workflow Description Language (GWorkflowDL)~\cite{Alt2006}.
GWES enables key workflow development and deployment lifecycle phases including
visual composition, data-staging, Grid middleware interfacing and execution
monitoring. Currently, GWES is intensively being used in the field of Medical
Image Processing especially at the Charit\'{e} Hospital at Berlin in Germany.
It has been successfully interfaced with the German D-Grid infrastructure.
%

@@

\begin{table}[ht!]
\begin{center}
% use packages: array
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
\textbf{SWE} / & \multicolumn{3}{c|}{\textbf{Parallelism}} & \multicolumn{3}{c|}{\textbf{Optimization}}\\
\cline{2-7}
\textbf{Language} & \textbf{data} & \textbf{workflow} & \textbf{pipeline} & \textbf{static} & \textbf{runtime} & \textbf{other}\\ \hline
Taverna/Scufl2 & \checkmark & \checkmark & \checkmark & -- & \checkmark & -- \\\hline
+/BPEL &foreach & \checkmark & -- &-- &\checkmark &-- \\\hline
Triana/- & \checkmark & \checkmark & \checkmark & -- & \checkmark &--\\\hline
MOTEUR/Scufl & \checkmark & \checkmark & \checkmark &-- &\checkmark &-- \\\hline
Kepler/MoML & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &--\\\hline
GridNexus/GXPL & -- & \checkmark & -- &-- &\checkmark &-- \\\hline
Swift/SwiftScript & \checkmark & \checkmark & \checkmark & \checkmark &\checkmark &-- \\\hline
Askalon/AGWL &--& \checkmark & -- & \checkmark & \checkmark & -- \\\hline
P-GRADE/- &\checkmark &\checkmark &\checkmark &--&--&--\\\hline
Vizbuilder/bioflow &-- &\checkmark &-- &--&\checkmark &Query optimization\\\hline
Martlet/- & \checkmark & \checkmark & -- &-- &\checkmark&-- \\\hline
Knime/- & \checkmark &\checkmark &\checkmark & -- & \checkmark & -- \\\hline
StarFlow/- &\checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -- \\\hline
Pegasus/DAX &\checkmark &\checkmark &\checkmark &\checkmark &\checkmark &AI-based techniques \\\hline
Galaxy/- &-- &-- &\checkmark &-- &-- &-- \\\hline
GWES/GWorkflowDL &-- &-- &\checkmark &-- &\checkmark &-- \\\hline
MOTEUR2/Gwendia~ & \checkmark &\checkmark & \checkmark &-- &\checkmark &-- \\\hline
\end{tabular}
%\caption{A summary of SWEs Scalability and optimization Mechanisms}
\label{tbl:scalability}
\end{center}
\end{table}

@@

\begin{table}[ht!]
\begin{center}
\hspace*{-2.2cm}
% use packages: array
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\textbf{SWE /} & \multicolumn{3}{c|}{\textbf{Language Level}} & \multicolumn{2}{c|}{\textbf{Resource Level}}\\
\cline{2-6}
\multirow{2}{*}{\textbf{Language}} & \textbf{Data Coupling} & \textbf{Data Types} & \textbf{Data Structures} & \textbf{Type} & \textbf{Access}\\
& \textbf{with Workflow} & \textbf{Supported} & \textbf{Supported} & & \\
\hline
Taverna/Scufl2&Loose&MIME-types&multi-dim Arrays&Local+Remote&By Reference\\\hline
+/BPEL&Strong&XSD-Based&--&Grids, Clouds& Adhoc\\\hline
%
\multirow{2}{*}{Triana/-}&Strong&XML Schema-&Streams,&Local+&Through \\
&&Types&Java Objects&Remote&Adaptors\\\hline
%
\multirow{2}{*}{MOTEUR/Scufl}&Loose&MIME-&Lists&Local+&Services+\\
&&Types&&Remote&Middleware\\\hline
% 
\multirow{2}{*}{Kepler/}&Strong&Primitive&Collections,&Filesystem&Customized\\
MoML&&Types&User defined&&Actors\\\hline
% 
GridNexus/GXPL&Loose&XSD-Based&--&Grid Data&OGSA-DAI \\\hline
% 
\multirow{2}{*}{Swift/SwiftScript} &Strong&--&1-Dim Arrays,&Shared& POSIX-style\\
&&&Customised&FileSystem&\\\hline
% 
Askalon/AGWL&Loose&MIME-types&Arrays&Data Repository&SQL-based\\\hline
%
\multirow{2}{*}{P-GRADE/-}&Loose&--&--&Grid&Secure\\
&&&&File Catalogs&Protocols\\\hline
%
Vizbuilder/bioflow&Strong&SQL-types&--&Database&SQL,PL/SQL\\\hline
%
\multirow{2}{*}{Martlet/-}&Loose&URI&Lists&Distributed& By\\
&&&&Filesystems&Reference\\\hline
% 
\multirow{2}{*}{Knime/-}&Loose&String, Integer,&Filestreams,&FileSystems,&Datamining\\
&&Double&Arrays&Databases&Tools\\\hline
% 
StarFlow/-&Strong&Python-based&Python-based&--&--\\\hline
GWES/GWorkflowDL&Strong&java-based&java-based&--&--\\\hline
% 
\multirow{2}{*}{}Pegasus/&Loose&--&--&Distributed&Secure\\
DAX&&&&Filesystems&Protocols\\\hline
% 
Galaxy/-&Loose&--&File Streams&File Systems&--\\\hline
% 
\multirow{2}{*}{MOTEUR2/} &Loose&Integer,String,&multi-dim Arrays&Grid&By\\
Gwendia~ &&Double,URI&&Data&Reference\\\hline
\end{tabular}
%\caption{A summary of SWEs Data Management Mechanisms}
\label{tbl:datadesc}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
%\hspace*{-1cm}
% use packages: array
\begin{tabular}{|l|l|l|}
\hline
\textbf{SWE/Language} & \textbf{DCI Interfacing Mechanisms} & \textbf{Currently Interfaced DCIs}\\
\hline
Taverna/Scufl2 & By Extension & None Inherently \\\hline
+/BPEL & Ad-hoc & None Inherently \\\hline
Triana/- & Provides Adapters & P2P Networks, NGS (ngs.org.uk)  \\\hline
MOTEUR/Scufl & Inherent & EGI (www.eu-egi.eu), g5k (www.grid5000.fr)\\\hline
Kepler/MoML & Using Directors & EarthGrid (http://seek.ecoinformatics.org) \\\hline
GridNexus/GXPL & Inherent & Generic Grids \\\hline
Swift/SwiftScript & Inherent & IBM/bluegene/P HPC Clusters \\\hline
Askalon/AGWL & Inherent & Austrian Grid (http://www.austriangrid.at/)\\\hline
P-GRADE/- & Inherent & NGS \\\hline
Vizbuilder/bioflow & DCI Independent & None Inherently \\\hline
Martlet/- & interfaces multiple middlewares & Generic Grids \\\hline
Knime/- & Extensions & SGE Clusters \\\hline
StarFlow/- &Interfacing External Schedulars & Generic Clouds \\\hline
GWES/GWorkflowDL&Globus WS-Gram& German D-Grid \\\hline
Pegasus/DAX & Using Condor & OSG (www.opensciencegrid.org) \\\hline
Galaxy/- & Extensions & None Inherently \\\hline
MOTEUR2/Gwendia~ & Inherent & EGI, g5k, HIPernet Cloud \\\hline
\end{tabular}
%\vspace*{1cm}
%\caption{A summary of SWEs DCI Interfacing Mechanisms}
\label{tbl:interfacedci}
\end{center}
\end{table}

@@

%
\begin{table}
\begin{center}
% use packages: array
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{SWE/Language} &\textbf{Control Structures}& \textbf{Array Programming} & \textbf{Iterators}\\
\hline
Taverna/Scufl2 &Conditionals and Loops& \checkmark & \checkmark \\\hline
+/BPEL &Conditionals and Loops& -- & -- \\\hline
Triana/- &Conditionals and Loops& \checkmark & -- \\\hline
MOTEUR/Scufl &Conditionals& -- & \checkmark \\\hline
Kepler/MoML &Conditionals and Loops & \checkmark & limited \\\hline
GridNexus/GXPL &Conditionals and Loops& -- & -- \\\hline
Swift/SwiftScript & Conditionals and Loops & 1-dim & limited \\\hline
Askalon/AGWL & Conditionals and Loops & -- & -- \\\hline
P-GRADE/- & Conditionals & -- & \checkmark \\\hline
Vizbuilder/bioflow & Conditionals and Loops & -- & -- \\\hline
Martlet/- & Conditionals and Loops & -- & -- \\\hline
Knime/- & Loops & -- & \checkmark \\\hline
StarFlow/- & Conditionals and Loops & -- & -- \\\hline
GWES/GWorkflowDL & Conditionals& -- & -- \\\hline
Pegasus/DAX & -- & \checkmark & -- \\\hline
Galaxy/- & -- & -- & -- \\\hline
MOTEUR2/Gwendia~ & Conditionals and Loops & \checkmark & \checkmark \\\hline
\end{tabular}
%\caption{A summary of SWEs Expressiveness Capabilities}
\label{tbl:expressive}
\end{center}
\end{table}

@@

\begin{table}[ht!]
\begin{center}
% use packages: array
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{SWE/Language} & \textbf{Workflow Composition} & \textbf{Reusable} & \textbf{Extendable}\\
\hline
Taverna/Scufl2 & GUI & \checkmark & \checkmark\\\hline
+/BPEL & GUI & \checkmark & \checkmark \\\hline
Triana/- & GUI & \checkmark & \checkmark \\\hline
MOTEUR/Scufl & GUI & \checkmark & -- \\\hline
Kepler/MoML & GUI & \checkmark & \checkmark \\\hline
GridNexus/GXPL & GUI/Text & -- & -- \\\hline
Swift/SwiftScript & Text & -- & -- \\\hline
Askalon/AGWL & GUI & \checkmark & \checkmark \\\hline
P-GRADE/- & Web Portal & \checkmark & -- \\\hline
Vizbuilder/bioflow & GUI/Text & \checkmark & -- \\\hline
Martlet/- & Text & -- & -- \\\hline
Knime/- & GUI & \checkmark & -- \\\hline
StarFlow/- & Text & \checkmark & -- \\\hline
GWES/GWorkflowDL & GUI & \checkmark & \checkmark \\\hline
Pegasus/DAX & GUI & \checkmark & \checkmark \\\hline
Galaxy/- & Web Portal & -- & -- \\\hline
\multirow{2}{*}{MOTEUR2/} & GUI/Text & \checkmark & --\\
Gwendia~/gscript & & & \\\hline
\end{tabular}
%\caption{A summary of SWEs Usability Analysis}
\label{tbl:usability}
\end{center}
\end{table}

@@

Business Process Execution Language (BPEL) \cite{oasis:2007} is a workflow
language pioneered by IBM Corporation and used predominantly to define business
workflows. It is now a broad standard for business workflow applications. BPEL
is an open language and commercial enactors are available that provide
enactment functionality for the BPEL workflows. There has been attempts to use
BPEL or its extended version in the scientific applications
\cite{slominski:2007,bpel}. However, the language does not suit well to the
data-intensive, grid workflows and has a limited adoption in the scientific
community
\cite{tan-missier-etal:2010,barga-gannon:2007,scherp-hoing-etal:2009,emmerich-butchart-etal:2005}.
Apart from that, a lack of an open-source workflow engine with graphical
interface also contributes to its limited acceptance in academic research.
However, BPEL has been famously unsuitable for data intensive scientific
applications  BPEL lacks different user-defined strategies to effect multiple
invocations of each processor according to the application semantics. BPEL also
lacks data streams and implicit parallelism forcing a user to explicitly define
parallelism using one of its parallel processing constructs such as
\texttt{foreach}.

>>>>

%Workflows considered:
% MOTEUR2
% Taverna
% Swift
% Pegasus
% Triana
% Galaxy
% Makeflow
% Kepler/Ptolemy
% Dryad
% Science pipes
% Spark
% Linda
% DaGue

@@

\section{Objectives and Organization}
The present work analyzes contemporary SWEs in order to find the capabilities
and key features in terms of the characteristics described in the introduction
part of this paper. The rest of this paper is organized as follows: Section
\ref{sec:overview} presents an overview of the key SWEs studied in this work.
Section \ref{sec:scal} discusses the optimization features and resulting
scalable performance as observed in the SWEs studied. Section \ref{sec:data}
studies the data management and description aspects of the SWEs. Section
\ref{sec:inter} presents a study of SWEs approaches to interface the
large-scale computing infrastructures. Section \ref{sec:lang} details the
language features and expressiveness of the SWEs studied. Section
\ref{sec:usable} discusses the usability aspects of the SWEs.  Section
\ref{sec:sum} presents a summary and a classification of all the aspects of
SWEs studied in this work. Throughout the paper, we also present tables
summarizing the aspects of SWEs considered in respective sections. It was not
possible to obtain an exhaustive information about all the SWEs considered in
this work since we rely on the peer reviewed publications for information on a
given SWE. Some features, such as extendability of SWE strongly depend on the
extension points designed in the code; consequently, we consider it present in
only those SWEs that have a documented description or demonstration of the
given feature.

@@

% Put this somewhere else, either in the analysis or may be in the conclusions
% and summary.
%The \textit{gscript} \cite{maheshwari-montagnat:2010} workflow language
%developed as a part of the current work is similar in spirit to the
%SwiftScript. Similar to SwiftScript, gscript adopts futures to express
%asynchronous execution in a script-based language. A SwiftScript workflow
%contains statically typed datasets that can be specified using ``annotations''
%in the script. Unlike SwiftScript, gscript does provide user-defined
%strategies in order to effect data combinations in the workflow. SwiftScript
%provides for fault-tolerance and exception handling at run-time which is
%lacking currently in gscript. The underlying datasets organization in swift is
%represented using an XDTM XML-representation. This means that effectively, a
%workflow is SwiftScript plus the XDTM data representation. A workflow is
%simpler and compact in gscript as its data description is embedded within a
%workflow representation. SwiftScript supports creation of nested-workflows in
%the form of compound procedures. Conditional and Loop execution is supported
%using the if and for/while statements in both the languages.

@@

@@

\begin{table}[ht!]
\begin{center}
% use packages: array
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
\textbf{SWE} / & \multicolumn{3}{c|}{\textbf{Parallelism}} & \multicolumn{3}{c|}{\textbf{Optimization}}\\
\cline{2-7}
\textbf{Language} & \textbf{data} & \textbf{workflow} & \textbf{pipeline} & \textbf{static} & \textbf{runtime} & \textbf{other}\\ \hline
Taverna/Scufl2 & \checkmark & \checkmark & \checkmark & -- & \checkmark & -- \\\hline
+/BPEL &foreach & \checkmark & -- &-- &\checkmark &-- \\\hline
Triana/- & \checkmark & \checkmark & \checkmark & -- & \checkmark &--\\\hline
MOTEUR/Scufl & \checkmark & \checkmark & \checkmark &-- &\checkmark &-- \\\hline
Kepler/MoML & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &--\\\hline
GridNexus/GXPL & -- & \checkmark & -- &-- &\checkmark &-- \\\hline
Swift/SwiftScript & \checkmark & \checkmark & \checkmark & \checkmark &\checkmark &-- \\\hline
Askalon/AGWL &--& \checkmark & -- & \checkmark & \checkmark & -- \\\hline
P-GRADE/- &\checkmark &\checkmark &\checkmark &--&--&--\\\hline
Vizbuilder/bioflow &-- &\checkmark &-- &--&\checkmark &Query optimization\\\hline
Martlet/- & \checkmark & \checkmark & -- &-- &\checkmark&-- \\\hline
Knime/- & \checkmark &\checkmark &\checkmark & -- & \checkmark & -- \\\hline
StarFlow/- &\checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -- \\\hline
Pegasus/DAX &\checkmark &\checkmark &\checkmark &\checkmark &\checkmark &AI-based techniques \\\hline
Galaxy/- &-- &-- &\checkmark &-- &-- &-- \\\hline
MOTEUR2/Gwendia~ & \checkmark &\checkmark & \checkmark &-- &\checkmark &-- \\\hline
\end{tabular}
%\caption{A summary of SWEs Scalability and optimization Mechanisms}
\label{tbl:scalability}
\end{center}
\end{table}

@@

\begin{table}[ht!]
\begin{center}
\hspace*{-2.2cm}
% use packages: array
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\textbf{SWE /} & \multicolumn{3}{c|}{\textbf{Language Level}} & \multicolumn{2}{c|}{\textbf{Resource Level}}\\
\cline{2-6}
\multirow{2}{*}{\textbf{Language}} & \textbf{Data Coupling} & \textbf{Data Types} & \textbf{Data Structures} & \textbf{Type} & \textbf{Access}\\
& \textbf{with Workflow} & \textbf{Supported} & \textbf{Supported} & & \\
\hline
Taverna/Scufl2&Loose&MIME-types&multi-dim Arrays&Local+Remote&By Reference\\\hline
+/BPEL&Strong&XSD-Based&--&Grids, Clouds& Adhoc\\\hline
%
\multirow{2}{*}{Triana/-}&Strong&XML Schema-&Streams,&Local+&Through \\
&&Types&Java Objects&Remote&Adaptors\\\hline
%
\multirow{2}{*}{MOTEUR/Scufl}&Loose&MIME-&Lists&Local+&Services+\\
&&Types&&Remote&Middleware\\\hline
% 
\multirow{2}{*}{Kepler/}&Strong&Primitive&Collections,&Filesystem&Customized\\
MoML&&Types&User defined&&Actors\\\hline
% 
GridNexus/GXPL&Loose&XSD-Based&--&Grid Data&OGSA-DAI \\\hline
% 
\multirow{2}{*}{Swift/SwiftScript} &Strong&--&1-Dim Arrays,&Shared& POSIX-style\\
&&&Customised&FileSystem&\\\hline
% 
Askalon/AGWL&Loose&MIME-types&Arrays&Data Repository&SQL-based\\\hline
%
\multirow{2}{*}{P-GRADE/-}&Loose&--&--&Grid&Secure\\
&&&&File Catalogs&Protocols\\\hline
%
Vizbuilder/bioflow&Strong&SQL-types&--&Database&SQL,PL/SQL\\\hline
%
\multirow{2}{*}{Martlet/-}&Loose&URI&Lists&Distributed& By\\
&&&&Filesystems&Reference\\\hline
% 
\multirow{2}{*}{Knime/-}&Loose&String, Integer,&Filestreams,&FileSystems,&Datamining\\
&&Double&Arrays&Databases&Tools\\\hline
% 
StarFlow/-&Strong&Python-based&Python-based&--&--\\\hline
% 
\multirow{2}{*}{}Pegasus/&Loose&--&--&Distributed&Secure\\
DAX&&&&Filesystems&Protocols\\\hline
% 
Galaxy/-&Loose&--&File Streams&File Systems&--\\\hline
% 
\multirow{2}{*}{MOTEUR2/} &Loose&Integer,String,&multi-dim Arrays&Grid&By\\
Gwendia~ &&Double,URI&&Data&Reference\\\hline
\end{tabular}
%\caption{A summary of SWEs Data Management Mechanisms}
\label{tbl:datadesc}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
%\hspace*{-1cm}
% use packages: array
\begin{tabular}{|l|l|l|}
\hline
\textbf{SWE/Language} & \textbf{DCI Interfacing Mechanisms} & \textbf{Currently Interfaced DCIs}\\
\hline
Taverna/Scufl2 & By Extension & None Inherently \\\hline
+/BPEL & Ad-hoc & None Inherently \\\hline
Triana/- & Provides Adapters & P2P Networks, NGS (ngs.org.uk)  \\\hline
MOTEUR/Scufl & Inherent & EGI (www.eu-egi.eu), g5k (www.grid5000.fr)\\\hline
Kepler/MoML & Using Directors & EarthGrid (http://seek.ecoinformatics.org) \\\hline
GridNexus/GXPL & Inherent & Generic Grids \\\hline
Swift/SwiftScript & Inherent & IBM/bluegene/P HPC Clusters \\\hline
Askalon/AGWL & Inherent & Austrian Grid (http://www.austriangrid.at/)\\\hline
P-GRADE/- & Inherent & NGS \\\hline
Vizbuilder/bioflow & DCI Independent & None Inherently \\\hline
Martlet/- & interfaces multiple middlewares & Generic Grids \\\hline
Knime/- & Extensions & SGE Clusters \\\hline
StarFlow/- &Interfacing External Schedulars & Generic Clouds \\\hline
Pegasus/DAX & Using Condor & OSG (www.opensciencegrid.org) \\\hline
Galaxy/- & Extensions & None Inherently \\\hline
MOTEUR2/Gwendia~ & Inherent & EGI, g5k, HIPernet Cloud \\\hline
\end{tabular}
%\vspace*{1cm}
%\caption{A summary of SWEs DCI Interfacing Mechanisms}
\label{tbl:interfacedci}
\end{center}
\end{table}

@@

%
\begin{table}
\begin{center}
% use packages: array
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{SWE/Language} &\textbf{Control Structures}& \textbf{Array Programming} & \textbf{Iterators}\\
\hline
Taverna/Scufl2 &Conditionals and Loops& \checkmark & \checkmark \\\hline
+/BPEL &Conditionals and Loops& -- & -- \\\hline
Triana/- &Conditionals and Loops& \checkmark & -- \\\hline
MOTEUR/Scufl &Conditionals& -- & \checkmark \\\hline
Kepler/MoML &Conditionals and Loops & \checkmark & limited \\\hline
GridNexus/GXPL &Conditionals and Loops& -- & -- \\\hline
Swift/SwiftScript & Conditionals and Loops & 1-dim & limited \\\hline
Askalon/AGWL & Conditionals and Loops & -- & -- \\\hline
P-GRADE/- & Conditionals & -- & \checkmark \\\hline
Vizbuilder/bioflow & Conditionals and Loops & -- & -- \\\hline
Martlet/- & Conditionals and Loops & -- & -- \\\hline
Knime/- & Loops & -- & \checkmark \\\hline
StarFlow/- & Conditionals and Loops & -- & -- \\\hline
Pegasus/DAX & -- & \checkmark & -- \\\hline
Galaxy/- & -- & -- & -- \\\hline
MOTEUR2/Gwendia~ & Conditionals and Loops & \checkmark & \checkmark \\\hline
\end{tabular}
%\caption{A summary of SWEs Expressiveness Capabilities}
\label{tbl:expressive}
\end{center}
\end{table}

@@

\begin{table}[ht!]
\begin{center}
% use packages: array
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{SWE/Language} & \textbf{Workflow Composition} & \textbf{Reusable} & \textbf{Extendable}\\
\hline
Taverna/Scufl2 & GUI & \checkmark & \checkmark\\\hline
+/BPEL & GUI & \checkmark & \checkmark \\\hline
Triana/- & GUI & \checkmark & \checkmark \\\hline
MOTEUR/Scufl & GUI & \checkmark & -- \\\hline
Kepler/MoML & GUI & \checkmark & \checkmark \\\hline
GridNexus/GXPL & GUI/Text & -- & -- \\\hline
Swift/SwiftScript & Text & -- & -- \\\hline
Askalon/AGWL & GUI & \checkmark & \checkmark \\\hline
P-GRADE/- & Web Portal & \checkmark & -- \\\hline
Vizbuilder/bioflow & GUI/Text & \checkmark & -- \\\hline
Martlet/- & Text & -- & -- \\\hline
Knime/- & GUI & \checkmark & -- \\\hline
StarFlow/- & Text & \checkmark & -- \\\hline
Pegasus/DAX & GUI & \checkmark & \checkmark \\\hline
Galaxy/- & Web Portal & -- & -- \\\hline
\multirow{2}{*}{MOTEUR2/} & GUI/Text & \checkmark & --\\
Gwendia~/gscript & & & \\\hline
\end{tabular}
%\caption{A summary of SWEs Usability Analysis}
\label{tbl:usability}
\end{center}
\end{table}

@@

Business Process Execution Language (BPEL) \cite{oasis:2007} is a workflow
language pioneered by IBM Corporation and used predominantly to define business
workflows. It is now a broad standard for business workflow applications. BPEL
is an open language and commercial enactors are available that provide
enactment functionality for the BPEL workflows. There has been attempts to use
BPEL or its extended version in the scientific applications
\cite{slominski:2007,bpel}. However, the language does not suit well to the
data-intensive, grid workflows and has a limited adoption in the scientific
community
\cite{tan-missier-etal:2010,barga-gannon:2007,scherp-hoing-etal:2009,emmerich-butchart-etal:2005}.
Apart from that, a lack of an open-source workflow engine with graphical
interface also contributes to its limited acceptance in academic research.
However, BPEL has been famously unsuitable for data intensive scientific
applications  BPEL lacks different user-defined strategies to effect multiple
invocations of each processor according to the application semantics. BPEL also
lacks data streams and implicit parallelism forcing a user to explicitly define
parallelism using one of its parallel processing constructs such as
\texttt{foreach}.
%
\paragraph{GWES/GWorkflowDL} The Grid Workflow Execution Service
(GWES)~\cite{Hoheisel2007} is a SWE developed by the Fraunhofer-First
institute. It implements a dynamic workflow composition and enactment model
based on the Grid Workflow Description Language (GWorkflowDL)~\cite{Alt2006}.
GWES enables key workflow development and deployment lifecycle phases including
visual composition, data-staging, Grid middleware interfacing and execution
monitoring. Currently, GWES is intensively being used in the field of Medical
Image Processing especially at the Charit\'{e} Hospital at Berlin in Germany.
It has been successfully interfaced with the German D-Grid infrastructure.
%
